Manuscript

### 2019-01-24 manuscript_analysis.Rmd
updated analyses in Rmd

```{r}
ccle_data_ptmuts <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes_ptmuts.rds")
ctrp_data_ptmuts <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes_ptmuts.rds")
gdsc_data_ptmuts <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes_ptmuts.rds")

@@

```{r}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
# crispr_data_ptmuts <- readRDS("./../htscreens_giant_files/crispr_data_18Q4_ptmuts.rds")
# 
# crispr_data_ptmuts <- filter(crispr_data, Hugo_Symbol %in% g2p_genes)
# crispr_data_ptmuts <- filter(crispr_data_ptmuts, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
# crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele == 1 | is.na(crispr_data_ptmuts$Reference_Allele), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
# crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)
# crispr_data_ptmuts <- subset(crispr_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
# crispr_data_ptmuts <- unique(crispr_data_ptmuts)

# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p$Mutation_Status_Nonsilent <- factor(crispr_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
```

@@

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
system.time({ ccle_data_allgenes <- WilcoxonByDrugAllGenes(ccle_melt, ccle_grid, "CCLE") })
saveRDS(ccle_data_allgenes, "./data_munging/rds/ccle_data_18Q3_allgenes.rds", compress = "xz")

@@

```{r}
test <- p.adjust(ccle_signif_g2p$p, method = "BH")
```

@@

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.var  s = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
ctrp_data_allgenes <- WilcoxonByDrugAllGenes(ctrp_melt, ctrp_grid, "CTRP")
saveRDS(ctrp_data_allgenes, "./data_munging/rds/ctrp_data_18Q3_allgenes.rds", compress = "xz")
```

@@

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
system.time({ gdsc_data_allgenes <- WilcoxonByDrugAllGenes(gdsc_melt, gdsc_grid, "GDSC") })
saveRDS(gdsc_data_allgenes, "./data_munging/rds/gdsc_data_18Q3_allgenes.rds", compress = "xz")
```

@@

```{r}
crispr_pearson_res_GE <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                         summarize(p = cor.test(y = Score, x = RPKM, method = "pearson", use = "complete.obs")$p.value,
                                   corr = cor.test(y = Score, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                                   Metric = "Gene Expression")
crispr_pearson_res_CN <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                         summarize(p = cor.test(y = Score, x = CN, method = "pearson", use = "complete.obs")$p.value,
                                   corr = cor.test(y = Score, x = CN, method = "pearson", use = "complete.obs")$estimate,
                                   Metric = "Copy Number")
crispr_spearman_res_GE <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                          summarize(p = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$p.value,
                                    corr = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                                    Metric = "Gene Expression")
crispr_spearman_res_CN <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                          summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value,
                                    corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                                    Metric = "Copy Number")
```

@@

```{r}
g2p_druggene_tally <- g2p_filt %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

@@

# ggsave("./plots/manuscript/crispr_venn.pdf", crispr_venn)
```

@@

Spearman correlations for gene expression and copy number:
```{r fig.height=6, fig.width=12}
crispr_spearman_res_GE <- transform(crispr_spearman_res_GE, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_GE$p_log10 <- -log10(crispr_spearman_res_GE$p)
crispr_spearman_res_CN <- transform(crispr_spearman_res_CN, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_CN$p_log10 <- -log10(crispr_spearman_res_CN$p)
spearman_KRAS <- rbind(filter(crispr_data_g2p, Hugo_Symbol == "KRAS") %>% summarize(p = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), filter(crispr_data_g2p, Hugo_Symbol == "KRAS") %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"))

@@

crispr_ge_spearman_corr_plot <- ggplot(data = crispr_spearman_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "deepskyblue4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)") +
  annotation_custom(ggplotGrob(crispr_ge_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 3.5, y = -0.285, xend = 13, yend = -0.16), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

crispr_cn_spearman_corr_plot <- ggplot(data = crispr_spearman_res_CN, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "green4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)") +
  annotation_custom(ggplotGrob(crispr_cn_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 2.5, y = -0.232, xend = 13, yend = -0.16), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

crispr_spearman_corr_plot <- plot_grid(crispr_ge_spearman_corr_plot, crispr_cn_spearman_corr_plot, align = "h")
crispr_spearman_corr_plot
# ggsave("./plots/manuscript/crispr_spearman_corr_plot.png", crispr_spearman_corr_plot, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

@@

ccle_summ_res_mut_plot <- ggplot(data = subset(ccle_res_mut, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), labels = seq(0, 35, by = 5)) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, label = Drug_Gene, color = Metric), force = 10, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  theme(legend.position = "none", panel.grid.major.x = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(5.5, 1.5, 5.5, 5.5), "pt"), strip.background.y = element_blank(), strip.text.y = element_blank()) +
  scale_color_manual(values = c("Mutation Status" = "darkorchid4")) +
  labs(y = "-log10(p-value)")

ccle_summ_res_pearson_plot <- ggplot(data = subset(ccle_pearson_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)")

@@

ccle_summ_res_spearman_plot <- ggplot(data = subset(ccle_spearman_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

@@

gdsc_summ_res_spearman_plot <- ggplot(data = subset(gdsc_spearman_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 50, by = 5), labels = seq(0, 50, by = 5)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 20, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

@@

all_spearman_res <- rbind(ccle_spearman_res_filt, ctrp_spearman_res_filt, gdsc_spearman_res_filt)
all_spearman_summ_plot <- ggplot(data = subset(all_spearman_res, p < 0.05)) +
  facet_grid(Dataset ~ Metric, scales = "free", space = "free") +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Dataset), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 60, by = 10), labels = seq(0, 60, by = 10)) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Dataset, label = Drug_Gene), force = 10, nudge_y = 1, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = "none", plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)", y = "-log10(p-value)")
all_spearman_summ_plot
# ggsave("./plots/manuscript/all_spearman_summ_plot.png", all_spearman_summ_plot, device = "png", dpi = 450, width = 12, height = 10, units = "in")
```

@@

ctrp_grid_summary_ge_corr <- cor.test(y = ctrp_grid_summary_data$AUC, x = ctrp_grid_summary_data$RPKM, method = "pearson", use = "complete.obs")
ctrp_grid_summary_ge_corr_text <- paste0("r(", ctrp_grid_summary_ge_corr$parameter, ") = ", round(ctrp_grid_summary_ge_corr$estimate, 2), "\np ", ifelse(round(ctrp_grid_summary_ge_corr$p.value, 3) == 0, "< 0.001", paste0("= ", round(ctrp_grid_summary_ge_corr$p.value, 3))))

@@

```{r}
ccle_data_g2p$G2P_DGL <- ifelse(ccle_data_g2p$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
ccle_data_g2p_split <- split(ccle_data_g2p, f = ccle_data_g2p$Drug_Gene)
ccle_corr_all <- MakeCorrResSummaryTable(ccle_data_g2p_split, dataset_name = "CCLE")

@@

# Comopare pancancer vs. lineage

@@

```{r}
test <- filter(compare_mut, Drug_Gene == "nilotinib_ABL1" & Dataset == "CCLE" & group_general_lineage_name == "bone cancer")
```

>>>>

```{r}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes.rds")
ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes.rds")
gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes.rds")
ccle_data_ptmuts <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes_ptmuts.rds")
ctrp_data_ptmuts <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes_ptmuts.rds")
gdsc_data_ptmuts <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes_ptmuts.rds")

@@

```{r}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
# Run next two lines only if the file was made (see end of the last chunk)
# crispr_data_ptmuts <- readRDS("./../htscreens_giant_files/crispr_data_18Q4_ptmuts.rds")
# crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)

# Filter CRISPR for G2P genes in G2P
crispr_data_ptmuts <- filter(crispr_data, Hugo_Symbol %in% g2p_genes)
# Filter G2P CRISPR for point mutations
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele == 1 | is.na(crispr_data_ptmuts$Reference_Allele), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)
crispr_data_ptmuts <- subset(crispr_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
crispr_data_g2p <- unique(crispr_data_ptmuts)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p$Mutation_Status_Nonsilent <- factor(crispr_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
```

@@

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
# ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
# ccle <- merge(drug_meta, ccle, by = "CID")
# ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
# ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
# system.time({ ccle_data_allgenes <- WilcoxonByDrugAllGenes(ccle_melt, ccle_grid, "CCLE") })
# saveRDS(ccle_data_allgenes, "./data_munging/rds/ccle_data_18Q3_allgenes.rds", compress = "xz")

@@

@@

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
# ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
# ctrp <- merge(drug_meta, ctrp, by = "CID")
# ctrp_melt <- melt(ctrp, id.var  s = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
# ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
# ctrp_data_allgenes <- WilcoxonByDrugAllGenes(ctrp_melt, ctrp_grid, "CTRP")
# saveRDS(ctrp_data_allgenes, "./data_munging/rds/ctrp_data_18Q3_allgenes.rds", compress = "xz")
```

@@

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
# gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
# gdsc <- merge(drug_meta, gdsc, by = "CID")
# gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
# gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
# system.time({ gdsc_data_allgenes <- WilcoxonByDrugAllGenes(gdsc_melt, gdsc_grid, "GDSC") })
# saveRDS(gdsc_data_allgenes, "./data_munging/rds/gdsc_data_18Q3_allgenes.rds", compress = "xz")
```

@@

```{r, warning=FALSE}
crispr_pearson_res_GE <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                         summarize(p = cor.test(y = Score, x = RPKM, method = "pearson", use = "complete.obs")$p.value,
                                   corr = cor.test(y = Score, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                                   Metric = "Gene Expression")
crispr_pearson_res_CN <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                         summarize(p = cor.test(y = Score, x = CN, method = "pearson", use = "complete.obs")$p.value,
                                   corr = cor.test(y = Score, x = CN, method = "pearson", use = "complete.obs")$estimate,
                                   Metric = "Copy Number")
crispr_spearman_res_GE <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                          summarize(p = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$p.value,
                                    corr = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                                    Metric = "Gene Expression")
crispr_spearman_res_CN <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                          summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value,
                                    corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                                    Metric = "Copy Number")
```

@@

```{r, warning=FALSE}
g2p_druggene_tally <- g2p_filt %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

@@

ggsave("./plots/manuscript/crispr_venn.pdf", crispr_venn)
```

@@

Spearman correlations for gene expression and copy number:
```{r fig.height=8, fig.width=16}
crispr_spearman_res_GE <- transform(crispr_spearman_res_GE, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_GE$p_log10 <- -log10(crispr_spearman_res_GE$p)
crispr_spearman_res_CN <- transform(crispr_spearman_res_CN, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_CN$p_log10 <- -log10(crispr_spearman_res_CN$p)
spearman_KRAS <- rbind(filter(crispr_data_g2p, Hugo_Symbol == "KRAS") %>% summarize(p = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), filter(crispr_data_g2p, Hugo_Symbol == "KRAS") %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"))

@@

crispr_ge_spearman_corr_plot <- ggplot(data = crispr_spearman_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "deepskyblue4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)") +
  annotation_custom(ggplotGrob(crispr_ge_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 3.5, y = -0.27, xend = 14, yend = -0.165), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

crispr_cn_spearman_corr_plot <- ggplot(data = crispr_spearman_res_CN, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "green4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)") +
  annotation_custom(ggplotGrob(crispr_cn_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 2.5, y = -0.232, xend = 14, yend = -0.165), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

crispr_spearman_corr_plot <- plot_grid(crispr_ge_spearman_corr_plot, crispr_cn_spearman_corr_plot, align = "h")
crispr_spearman_corr_plot
# ggsave("./plots/manuscript/crispr_spearman_corr_plot.png", crispr_spearman_corr_plot, device = "png", dpi = 450, width = 16, height = 8, units = "in")
```

@@

ccle_summ_res_mut_plot <- ggplot(data = subset(ccle_res_mut, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), labels = seq(0, 35, by = 5)) +
  coord_cartesian(ylim = c(0, 25)) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, label = Drug_Gene, color = Metric), force = 10, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  theme(legend.position = "none", panel.grid.major.x = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(5.5, 1.5, 5.5, 5.5), "pt"), strip.background.y = element_blank(), strip.text.y = element_blank()) +
  scale_color_manual(values = c("Mutation Status" = "darkorchid4")) +
  labs(y = "-log10(p-value)")

ccle_summ_res_pearson_plot <- ggplot(data = subset(ccle_pearson_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 25)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)")

@@

ccle_summ_res_spearman_plot <- ggplot(data = subset(ccle_spearman_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 25)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

@@

gdsc_summ_res_spearman_plot <- ggplot(data = subset(gdsc_spearman_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 30)) +
  scale_y_continuous(breaks = seq(0, 50, by = 5), labels = seq(0, 50, by = 5)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 20, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  annotate(geom = "text", x = 0, y = 30, label = c("afatinib_ERBB2 and lapatinib_ERBB2\n-log10(p-value) >> 100", ""), size = 3) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

@@

all_spearman_res <- rbind(ccle_spearman_res_filt, ctrp_spearman_res_filt, gdsc_spearman_res_filt)
all_spearman_summ_plot <- ggplot(data = subset(all_spearman_res, p < 0.05)) +
  facet_grid(Dataset ~ Metric, scales = "free", space = "free") +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Dataset), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 60, by = 10), labels = seq(0, 60, by = 10)) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Dataset, label = Drug_Gene), force = 10, nudge_y = 1, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = "none", plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  annotate(geom = "text", x = 0, y = 30, label = c("", "", "", "", "afatinib_ERBB2 and lapatinib_ERBB2\n-log10(p-value) >> 100", ""), size = 3) +
  labs(x = "Pearson correlation coefficient (r)", y = "-log10(p-value)")
all_spearman_summ_plot
# ggsave("./plots/manuscript/all_spearman_summ_plot.png", all_spearman_summ_plot, device = "png", dpi = 450, width = 12, height = 10, units = "in")
```

@@

ctrp_grid_summary_ge_corr <- cor.test(y = ctrp_grid_summary_data$AUC, x = ctrp_grid_summary_data$RPKM, method = "pearson", use = "complete.obs")
ctrp_grid_summary_ge_corr_text <- paste0("r(", ctrp_grid_summary_ge_corr$parameter, ") = ", formatC(ctrp_grid_summary_ge_corr$estimate, digits = 2, format = "f"), "\np ", ifelse(round(ctrp_grid_summary_ge_corr$p.value, 3) == 0, "< 0.001", paste0("= ", round(ctrp_grid_summary_ge_corr$p.value, 3))))

@@

```{r, warning=FALSE}
# cannot compute exact p-value with ties??????

ccle_data_g2p$G2P_DGL <- ifelse(ccle_data_g2p$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
ccle_data_g2p_split <- split(ccle_data_g2p, f = ccle_data_g2p$Drug_Gene)
ccle_corr_all <- MakeCorrResSummaryTable(ccle_data_g2p_split, dataset_name = "CCLE")

@@

# Compare pancancer vs. lineage

@@
